Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-03-06 src/tex/ms.tex
fix algorithm

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{afterpage}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tabularx,booktabs,multirow}
\usepackage{float}
\usepackage[italicdiff]{physics}

@@

With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
\begin{algorithm}[H]
	\hspace*{\algorithmicindent} \textbf{Input} $\mathcal{L}_{\mathrm{ini}}$
	\begin{algorithmic}[1]
		\State{$\mathcal{L}(\lambda) \gets \mathcal{L}_{\mathrm{ini}}$}
		\State{$\alpha \gets 10^{-6}$}
		\State{$i \gets 0$}
		\While{$i<30000$}
		\State{$\mathcal{L}'(\lambda) \gets \texttt{jax.grad}(\mathcal{L})$}
		\State{$\lambda \gets \lambda-\alpha\mathcal{L}'(\lambda)$}
		\EndWhile
	\end{algorithmic}
	\hspace*{\algorithmicindent} \textbf{Output} $\lambda$
	\caption{Gradient Descent}
\end{algorithm}
Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \citep{khan2016frequency}. We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find. With such an optimization procedure, we obtain the optimized coefficients.

@@

Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end with a lower mismatch. Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease. Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement. We see that in Fig.~\ref{fig:q148_fl}, the median of distribution has a $22.9\%$ decrease. However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}. This is due to the problem mentioned in section \ref{subsec:optimization}. Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.

@@

Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}. This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum instead of the constant spectrum.

@@

Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this feature since the original coefficients were fitted using NR waveforms with equal or similar spin, hence the model prefers waveforms with similar spin. One interesting feature is how the second and fourth quadrants respond to optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms generally improve with along optimization. However, mismatches in the fourth quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most waveforms even turned worse after optimization. These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}~and~\ref{fig:q148_fl}.

@@

In this paper, we have presented a systematic method to recalibrate GW models. This method utilizes {\jax}'s automatic differentiation to apply derivative-based optimization to recalibrate GW models jointly. Using the new implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in conjunction with NR waveforms from the SXS catalog, we recalibrate waveform coefficients of the IMRPhenomD model. In general, the waveform accuracy can be improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement. In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148} and \ref{fig:q148_fl}. By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.

>>>>

\usepackage{graphicx}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}

\SetKwInput{Parameters}{Parameters}
\SetKwInput{Variables}{Variables}

\usepackage{tabularx,booktabs,multirow}
\usepackage{float}
\usepackage[italicdiff]{physics}

@@

With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
\begin{algorithm}
	\caption{Gradient descent pseudocode}
	\KwIn{initial coefficients $\lambda_i$}
	\Parameters{number of iterations $N$, learning rate $\alpha$}
	\Variables{current coefficients $\lambda$, mismatch gradient $\nabla\mathcal{L}$}
	\KwResult{output coefficients $\lambda$}
	$\lambda\leftarrow\lambda_i$\\
	\tcc{Gradient Descent}
	\For{$i<N$}{
		$\mathcal{L}\leftarrow Mismatch(\lambda)$ \\
		$\nabla\mathcal{L}\leftarrow AutoDiff(\mathcal{L})$\\
		$\lambda\leftarrow\lambda-\alpha\nabla\mathcal{L}$\\
	}
	\Return{$\lambda$}
\end{algorithm}
Here, $\lambda_i$ are the original coefficients given in \citep{khan2016frequency}. We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find. With such an optimization procedure, we obtain the optimized coefficients.

@@

Using a constant noise spectrum with $\mathcal{L}_{\mathrm{mean}}$, we compute the mismatch for all testing waveforms and plot the distribution of the mismatch in Fig. \ref{fig:q148}. We find the distribution's peak has shifted to the end with a lower mismatch. Quantitatively, the peak has lowered by almost an order-of-magnitude and the median of distribution has a $\sim50.0\%$ decrease. Using $\mathcal{L}_{\mathrm{fl}}$, the mismatch distribution also has a similar improvement. We see that the median of distribution has a $22.9\%$ decrease. However, the distribution does not have a clear peak as Fig.~\ref{fig:q148}. This is due to the problem mentioned in section \ref{subsec:optimization}. Nevertheless, both distributions show improvement and the potential problem did not affect the optimization scheme significantly.

@@

Similarly, using the same procedures, distributions of mismatches calculated using {\zdethp} noise spectrum show better improvement than the weighted mismatch. The shape of the distribution is similar to Fig.~\ref{fig:q148}. This effect is expected since the IMRPhenomD model was originally constructed and fitted using the {\zdethp} weighted mismatch. Thus, it should be a model that fits closely to the NR waveforms with the influence of {\zdethp} noise spectrum instead of the constant spectrum.

@@

Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this feature since the original coefficients were fitted using NR waveforms with equal or similar spin, hence the model prefers waveforms with similar spin. One interesting feature is how the second and fourth quadrants respond to optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms generally improve with along optimization. However, mismatches in the fourth quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most waveforms even turned worse after optimization. These waveforms correspond to the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.

@@

In this paper, we have presented a systematic method to recalibrate GW models. This method utilizes {\jax}'s automatic differentiation to apply derivative-based optimization to recalibrate GW models jointly. Using the new implementation of the IMRPhenomD model, {\ripple}, which is written in \jax, in conjunction with NR waveforms from the SXS catalog, we recalibrate waveform coefficients of the IMRPhenomD model. In general, the waveform accuracy can be improved by 50\%. Comparing {\zdethp} weighted and unweighted mismatch, weighted mismatches have a slightly better improvement. In contrast, different types of loss function result in significantly different final mismatch distributions, where the result can be seen in Fig.~\ref{fig:q148}. By increasing the number of training waveforms, we see a slight improvement increase in Fig.~\ref{fig:q148_q1248_compare}.
